An approach for cloud correction of operational humidity channels

### 2020-09-10 Article/ici_submm_cloudcorr.tex
Added references to existing text in Introduction.

Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special role. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep[e.g.][]{kakar1983retrieval}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitions imposed by clouds. To be able to also sense humidity
inside and below clouds weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part several sensors, such as
MHS (Microwave Humidity Sounder) and ATMS (Advanced Technology Microwave
Sounder ).

Precipiation and most dense clouds, particularly if found at high altitude, can
still afffect measured radiances around 183\,GHz. As the impact from the
hydrometeors is then dominated by scattering, the complexity of the analysis of
the data increases dramatically and there exists a need to identify such cases.
This is normally denoted as cloud filtering, to obtain data of ``clear sky''
character. Such filtering has been applied to derive climate records, and is
also essentail in studies of the agreement between observations and
simulations. Examples on cloud filtering methods used for these applicatons are
\todo{(refs)}. These methods arebased on the 183\,GHz data alone and rules on
the brightness temperatures of different channels, including the difference
between channels, are applied to remove cloud affected data.

The main motivation for introducing 183\,GHz channels in operational sensors is
numerical weather prediction (NWP). Usage of microwave data in NWP by
``all-sky'' assimilation is growing, but 183\,GHz data are still mainly used in
a clear sky fashion. The later is particularly true in NWP of regional scope,
all-sky assimilation of 183\,GHz at many national weather agencies will likely
not be reached in many years. The cloud filtering in NWP differs between
centres, but normally \dots

>>>>

Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitions imposed by clouds. To be able to also sense humidity
inside and below clouds weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS (Advanced Technology Microwave Sounder, \citet{weng2012introduction}).

Precipiation and most dense clouds, particularly if found at high altitude, can
still afffect measured radiances around 183\,GHz
\citep[e.g.][]{bennartz2003sensitivity}. As the impact from the hydrometeors
then is dominated by scattering, the complexity of the analysis of the data
increases dramatically and there exists a need to identify the problematic
cases. This is normally denoted as cloud filtering, to obtain data of ``clear
sky'' character. Such filtering has been applied to derive climate records
\citep{lang2020new} and is essentail in studies of the agreement between
observations and simulations \citep{brogniez2016review} as well as 
comparing observations of different instruments to validate their calibration
\citep{john2013assessment,moradi:retri:15,berg2016intercalibration}. A commonly
used cloud filtering methods for these applicatons is the one of
\citet{buehler:aclou:07}. This method is based on the 183\,GHz data alone,
involving rules on the brightness temperatures differences between channels. An
older version is \citet{burns1997effects}.

The main motivation for introducing 183\,GHz channels in operational sensors is
numerical weather prediction (NWP). Usage of microwave data by ``all-sky''
assimilation is growing \citep{geer2017growing}, but 183\,GHz data are still
mainly used in a clear sky fashion \citep{geer2018all}. The later is
particularly true in NWP of regional scope, all-sky assimilation of 183\,GHz at
many national weather agencies will likely not be reached in many years. The
cloud filtering in NWP differs between centres. One approach is to make use of
a ``scattering index'' \citep{bennartz2002precipitation}, based on the
observations alone. Today likely more common is to use ``observation minus
background'' (O-B), where the forecast model is used to obtain an estimate of
the expected clear-sky value and the observation is rejected if the deviation
exceeds some threshold (???). In both cases, the filtering typically involving
observations around 89 and/or 150\,GHz.

% favrichon2019detecting?
